---
title: "chemspiderapi"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- badges: start -->
[![CircleCI build status](https://circleci.com/gh/NIVANorge/chemspiderapi.svg?style=shield)](https://circleci.com/gh/NIVANorge/chemspiderapi)
[![CodeCov test coverage](https://codecov.io/gh/NIVANorge/chemspiderapi/branch/master/graph/badge.svg)](https://codecov.io/gh/NIVANorge/chemspiderapi?branch=master)
<!-- badges: end -->

> R functionalities for ChemSpider's new API services

ChemSpider has introduced a new API syntax in late 2018, and [the old ChemSpider API syntax will be shut down at the end of November 2018](http://link.rsc.org/rsps/m/xSq8Cm8ovjN8-Elm0eYB3Sey61zutqNIUUaMcyc14sQ). `chemspiderapi` provides R wrappers around the new API services from ChemSpider. 

The aim of this package is to:

1) Translate the new ChemSpider API services into R-friendly functions.
2) Include thorough quality checking *before* the query is send, to avoid using up the query quota on, e.g., spelling errors.
3) Implement the R functionality in a way that is suitable for both `base` and `tidyverse` programming.

`chemspiderapi` relies on API keys to access ChemSpider's API services. For this, we recommend storing the ChemSpider API key using the [`keyring`](https://cran.r-project.org/package=keyring) package.

To limit the rate of API queries, we recommend using the  [`ratelimitr`](https://cran.r-project.org/package=ratelimitr) package or `purrr::slowly()` within the  [`tidyverse`](https://cran.r-project.org/package=tidyverse) package collection.

To handle PNG images, the [`magick`](https://cran.r-project.org/package=magick) package is recommended.

We furthermore recommend the [`memoise`](https://cran.r-project.org/package=memoise) package to "remember" the results of API queries (i.e., to not ruin the API allowance).

## Installation

### R package

Install the package from GitHub (using the `remotes` package; automatically installed alongside [`devtools`](https://cran.r-project.org/package=devtools)):

```{r,eval=FALSE}
# install.packages("devtools")
remotes::install_github("NIVANorge/chemspiderapi")
```

The development version can be installed with:
```{r,eval=FALSE}
# install.packages("devtools")
remotes::install_github("NIVANorge/chemspiderapi", ref = "dev")
```


Currently the only tested continuous integration environment for `chemspiderapi` is Linux. It *should* install smoothly on Windows 10 and mac OS machines as well. Please open an issue if you run into any troubles. 

### Dependencies

`chemspiderapi` relies on two essential dependencies. The [`curl`](https://cran.r-project.org/package=curl) package is used to handle the API queries and the [`jsonlite`](https://cran.r-project.org/package=jsonlite) package is necessary to wrap and unwrap information for the queries.

If not already installed, these packages will be installed automatically when installing `chemspiderapi`. Should this result in trouble, the dependency packages can be installed manually:

```{r, eval=FALSE}
install.packages(c("curl", "jsonlite"))
```

If `curl` or `jsonlite` are missing, (almost) all functions of `chemspiderapi` will fail and throw an error.


## Coverage

As of `r Sys.Date()`, the following functionalities are implemented (`r signif((27 / 27) * 100, digits = 2)`% functionality with `r signif((27 / 27) * 100, digits = 2)`% annotation):

**FILTERING**

| ChemSpider Compound API                  | `chemspiderapi` Wrapper               | `chemspiderapi` Help File |
|:---------------------------------------- |:------------------------------------- |:-------------------------:|
| filter-element-post                      | `post_element()`                      | yes                       |
| filter-formula-batch-post                | `post_formula_batch()`                | yes                       |
| filter-formula-batch-queryId-results-get | `get_formula_batch_queryId_results()` | yes                       |
| filter-formula-batch-queryId-status-get  | `get_formula_batch_queryId_status()`  | yes                       |
| filter-formula-post                      | `post_formula()`                      | yes                       |
| filter-inchi-post                        | `post_inchi()`                        | yes                       |
| filter-inchikey-post                     | `post_inchikey()`                     | yes                       |
| filter-intrinsicproperty-post            | `post_intrinsicproperty()`            | yes                       |
| filter-mass-batch-post                   | `post_mass_batch()`                   | yes                       |
| filter-mass-batch-queryId-results-get    | `get_mass_batch_queryId_results()`    | yes                       |
| filter-mass-batch-queryId-status-get     | `get_mass_batch_queryId_status()`     | yes                       |
| filter-mass-post                         | `post_mass()`                         | yes                       |
| filter-name-post                         | `post_name()`                         | yes                       |
| filter-queryId-results-get               | `get_queryId_results()`               | yes                       |
| filter-queryId-results-sdf-get           | `get_queryId_results_sdf()`           | yes                       |
| filter-queryId-status-get                | `get_queryId_status()`                | yes                       |
| filter-smiles-post                       | `post_smiles()`                       | yes                       |


**LOOKUPS**

| ChemSpider Compound API | `chemspiderapi` Wrapper | `chemspiderapi` Help File |
|:----------------------- |:----------------------- |:-------------------------:|
| lookups-datasources-get | `get_datasources()`     | yes                       |

**RECORDS**

| ChemSpider Compound API                 | `chemspiderapi` Wrapper             | `chemspiderapi` Help File |
|:--------------------------------------- |:----------------------------------- |:-------------------------:|
| records-batch-post                      | `post_batch()`                      | yes                       |
| records-recordId-details-get            | `get_recordId_details()`            | yes                       |
| records-recordId-externalreferences-get | `get_recordId_externalreferences()` | yes                       |
| records-recordId-image-get              | `get_recordId_image()`              | yes                       |
| records-recordId-mol-get                | `get_recordId_mol()`                | yes                       |

**TOOLS**

| ChemSpider Compound API      | `chemspiderapi` Wrapper    | `chemspiderapi` Help File |
|:---------------------------- |:-------------------------- |:-------------------------:|
| tools-convert-post           | `post_convert()`           | yes                       |
| tools-validate-inchikey-post | `post_validate_inchikey()` | yes                       |


## Best practices for ChemSpider's Compound APIs

_This section will be updated with practical examples in the future._

The basic workflow order for the above **FILTERING** queries is: 

1) POST Query

2) GET Status

3) GET Results (after GET Status returns `"Complete"`)

In practice, this means the following possible workflows can be implemented (from left to right):

| POST Query                 | GET Status                           | GET Results                           |
:--------------------------- |:------------------------------------ |:------------------------------------- |
| `post_element()`           | `get_queryId_status()`               | `get_queryId_results()`               |
| `post_formula()`           | `get_queryId_status()`               | `get_queryId_results()`               |
| `post_formula_batch()`     | `get_formula_batch_queryId_status()` | `get_formula_batch_queryId_results()` |
| `post_inchi()`             | `get_queryId_status()`               | `get_queryId_results()`               |
| `post_inchikey()`          | `get_queryId_status()`               | `get_queryId_results()`               |
| `post_intrinsicproperty()` | `get_queryId_status()`               | `get_queryId_results()`               |
| `post_mass()`              | `get_queryId_status()`               | `get_queryId_results()`               |
| `post_mass_batch()`        | `get_mass_batch_queryId_status()`    | `get_mass_batch_queryId_results()`    |
| `post_mass()`              | `get_queryId_status()`               | `get_queryId_results()`               |
| `post_name()`              | `get_queryId_status()`               | `get_queryId_results()`               |
| `post_smiles()`            | `get_queryId_status()`               | `get_queryId_results()`               |

Typically, the result will be one or multiple ChemSpider IDs (`recordId`). They can be used as input into the above **RECORDS** queries, e.g., `get_recordId_details()`.

## Vignettes

As of `r Sys.Date()`, the following five vignettes are available:

* **Storing and Accessing API Keys**: A basic example on how to safely store and retrieve API keys using the `keyring` package.

* **Rate-Limiting API Queries**: Multiple examples on how to slow down API queries. `base`, `ratelimitr`, and `tidyverse` workflows are provided.

* **Remembering API Queries**: A basic example on how to "remember" the results of API queries using the `memoise` package. This can greatly reduce redundant API queries.

* **Saving MOL and SDF Files**: Examples on how to save MOL and SDF files as returned by `get_recordId_mol()` and `get_queryId_results_sdf()`.

* **Saving PNG Images**: Examples on how to save PNG files, as returned by `get_recordId_image()`, using functionalities of the `magick` package.

## Funding

This package was created at the [Norwegian Institute for Water Research (NIVA)](https://www.niva.no/en) in conjunction with [NIVA's Computational Toxicology Program (NCTP)](https://www.niva.no/en/projectweb/nctp) at NIVA's [Section for Ecotoxicology and Risk Assessment](https://www.niva.no/en/research/ecotoxicology_and_risk_assessment) and funded by [The Research Council of Norway (RCN)](https://www.forskningsradet.no/en/), project 268294: [Cumulative Hazard and Risk Assessment of Complex Mixtures and Multiple Stressors (MixRisk)](https://www.forskningsradet.no/prosjektbanken/#/project/NFR/268294/Sprak=en).

## License

MIT &copy; Raoul Wolf
