---
title: "Rate-Limiting API Queries"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Rate-Limiting_API_Queries}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Most users of ChemSpider's API services will have to apply rate limiting conditions. Such a rate limit could be, for example, "no more than 15 queries per minute". The following two paragraphs describe how to rate-limit any `chemspiderapi` function; both in `base` R (using `ratelimitr`) and `tidyverse` coding styles.^[Technically, using the `purrr` package would be sufficient, but in practice most users will likely use the entire `tidyverse`.]

## Installing necessary packages

To run both examples, the `ratelimitr` package and the `tidyverse` package collection need to be installed. The neccessary API key is retrieved from a stored keyring using the `keyring` package; see separate vignette "Storing and Accessing API Keys" for instructions and recomendations on how to handle system-wide keys.

```{r,eval=FALSE}
install.packages(c("ratelimitr", "tidyverse"))
```

### Loading demo data

To start working with rate-limiting, the miniature demo data from `chemspiderapi` will be used. The data set is immediately available once `chemspiderapi` is loaded.

```{r setup}
## Loading chemspiderapi
library(chemspiderapi)

## Looking at the demo data
demo_chemicals
```


### A `base` R example

In `base` R, the the functionalities from the `ratelimitr` package will be used to rate-limit `chemspiderapi` functions. These rate-limited functions can then be used in, for example, the `apply()` family . Before diving into the rate-limiting, the API key needs to be retrieved.

```{r}
## Retrieving the API key from the keyring package.
## Note that the keyring package is not loaded, but only the keyring::key_get() function is called.
apikey <- keyring::key_get(service = "ChemSpider API key", username = Sys.getenv("USERNAME"))
```

To use the rate-limiting functionalities, we need to load `ratelimitr`

```{r}
## Loading the ratelimitr package
library(ratelimitr)
```

we will now create a rate-limited version of the `post_inchikey()` function and use it to obtain a query ID.

```{r,eval=FALSE}
## In ratelimitr::limit_rate(), the ratelimitr::rate() function specifies the rate limit as calls per second. 
slow_post_inchikey <- limit_rate(post_inchikey, rate(n = 15, period = 60))

## sapply() can be used in this context because the result of each iteration is a single vector.
demo_chemicals$queryId <- sapply(X = demo_chemicals$InChIKey, FUN = function(x) slow_post_inchikey(inchikey = x, apikey = apikey))

demo_chemicals
```

We can now check the status of query with a rate-limited `get_queryId_status()` function.
```{r,eval=FALSE}
slow_get_queryId_status <- limit_rate(get_queryId_status, rate(n = 15, period = 60))

## As the result of the query is an array, lapply() is used.
demo_chemicals$status <- lapply(X = demo_chemicals$queryId, FUN = function(x) slow_get_queryId_status(queryId = x, apikey = apikey))

demo_chemicals
```

All queries need to report the status `Complete` before proceeding. If this is the case, the ChemSpider IDs for the queries can be retrieved using a rate-limited `get_queryId_results()` function.
```{r,eval=FALSE}
slow_get_queryId_results <- limit_rate(get_queryId_results, rate(n = 15, period = 60))

## mapply() is used because the input covers two columns
demo_chemicals$id <- mapply(FUN = function(x, y) slow_get_queryId_results(queryId = x, status = unlist(y)[1], apikey = apikey), x = demo_chemicals$queryId, y = demo_chemicals$status)

demo_chemicals
```

Finally, the details for the chemicals can be retrieved using a rate-limited `get_recordId_details()` function.
```{r,eval=FALSE}
slow_get_recordId_details <- limit_rate(get_recordId_details, rate(n = 15, period = 60))

demo_chemicals$details <- lapply(X = demo_chemicals$id, FUN = function(x) slow_get_recordId_details(recordId = x, apikey = apikey, id = FALSE))

demo_chemicals
```


### A `tidyverse` example

In the `tidyverse` ecosystem, all relevant `chemspiderapi` functions can be used in functional programming approaches using the `purrr::map()` family. `purrr::slowly()` can be used to limit the number of calls of `chemspiderapi` functions in a given time interval.

```{r,eval=FALSE}
## Loading the tidyverse package family
library(tidyverse)

## Note that the API key is already available from the previous example!

## Using purrr::slowly() and purrr::rate_delay() to rate-limit the post_inchikey() function.
## In this example, we introduce a pause of four seconds and use the verbose output (`quiet = FALSE`).
slowly_post_inchikey <- slowly(post_inchikey, rate = rate_delay(pause = 4), quiet = FALSE)
```

The `post_inchikey()` function is now rate-limited as `slowly_post_inchikey()`. We can now use it to POST the five InChIKeys of the demo data.

```{r,eval=FALSE}
tidyverse_demo <- demo_chemicals %>% 
  as_tibble() %>% 
  mutate(queryId = map_chr(InChIKey, ~ slowly_post_inchikey(inchikey = .x, apikey = apikey)))

tidyverse_demo
```

In the next step, the queryId status needs to be checked using the `get_queryId_status()` function. This function as well needs to be rate-limited into `slowly_get_queryId_status()` before use. 

```{r,eval=FALSE}
slowly_get_queryId_status <- slowly(get_queryId_status, rate = rate_delay(pause = 4), quiet = FALSE)

tidyverse_demo <- tidyverse_demo %>% 
  mutate(status = map(queryId, ~ slowly_get_queryId_status(queryId = .x, apikey = apikey))) %>% 
  unnest(status)

tidyverse_demo
```

Ideally, the status column only contains `Complete`. If this is the case, the results can be retrieved using a rate-limited `slowly_get_queryId_results()` function.

```{r,eval=FALSE}
slowly_get_queryId_results <- slowly(get_queryId_results, rate = rate_delay(pause = 4), quiet = FALSE)

tidyverse_demo <- tidyverse_demo %>% 
  mutate(id = map2_int(queryId, status, ~ slowly_get_queryId_results(queryId = .x, status = .y, apikey = apikey)))

tidyverse_demo
```

The id column contains the ChemSpider IDs for the five chemicals; or rather their InChIKeys. We can now obtain details for the chemicals using the rate-limited version of `get_recordId_details()`.

```{r,eval=FALSE}
slowly_get_recordId_details <- slowly(get_recordId_details, rate = rate_delay(pause = 4), quiet = FALSE)

tidyverse_demo <- tidyverse_demo %>% 
  mutate(details = map(id, ~ slowly_get_recordId_details(id = .x, apikey = apikey, id = FALSE))) %>% 
  unnest(details)

tidyverse_demo
```
