---
title: "Rate-Limiting API Queries"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Rate-Limiting_API_Queries}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(chemspiderapi)
```

## Rate-limiting API queries for `chemspiderapi`

Most users of ChemSpider's API services will have to apply rate limiting conditions. Such a rate limit could be, for example, "no more than 15 queries per minute". The following two paragraph describe how to rate-limit any `chemspiderapi` function; both in `base` R (using `ratelimitr`) and `tidyverse` coding styles.

### A `base` R example

### A `tidyverse` example

In the tidyverse ecosystem, all relevant chemspiderapi functions can be used in functional programming approaches using the purrr::map() family of functions. purrr::slowly() can be used to limit the number of calls of chemspiderapi functions in a given time interval.

```{r,eval=FALSE}
## Installing the necessary packages if necessary
# install.packages(c("chemspiderapi", "tidyverse"))

## Loading the tidyverse package family and chemspiderapi
library(tidyverse)
library(keyring)
library(chemspiderapi)

## Storing the API key as temporary object via the keyring package
## See separate vignette for explanations on how to store security tokens using keyring

apikey <- keyring::key_get(service = "ChemSpider API key", username = Sys.getenv("LOGNAME"))

slowly_post_inchikey <- purrr::slowly(chemspiderapi::post_inchikey, rate = purrr::rate_delay(pause = 4), quiet = FALSE)

tidyverse_rl <- demo_data %>% 
  dplyr::select(InChIKey) %>% 
  dplyr::mutate(queryId = purrr::map_chr(InChIKey, ~ slowly_post_inchikey(inchikey = .x, apikey = apikey)))

tidyverse_rl
```

```{r,eval=FALSE}
slowly_get_queryId_status <- purrr::slowly(chemspiderapi::get_queryId_status, rate = rate_delay(pause = 4), quiet = FALSE)

tidyverse_rl <- tidyverse_rl %>% 
  dplyr::mutate(status = purrr::map(queryId, ~ slowly_get_queryId_status(queryId = .x, apikey = apikey)))

tidyverse_rl %>% 
  tidyr::unnest(status)

tidyverse_rl <- tidyverse_rl %>% 
  tidyr::unnest(status) %>% 
  dplyr::select(InChIKey, queryId, status)
```

```{r,eval=FALSE}
slowly_get_queryId_results <- slowly(chemspiderapi::get_queryId_results, rate = rate_delay(pause = 4), quiet = FALSE)

tidyverse_rl <- tidyverse_rl %>% 
  dplyr::mutate(id = purrr::map2_int(queryId, status, ~ slowly_get_queryId_results(queryId = .x, status = .y, apikey = apikey)))

tidyverse_rl
```

```{r,eval=FALSE}
slowly_get_recordId_details <- purrr::slowly(get_recordId_details, rate = rate_delay(pause = 4), quiet = FALSE)

tidyverse_rl <- tidyverse_rl %>% 
  dplyr::mutate(details = purrr::map(id, ~ slowly_get_recordId_details(id = .x, apikey = apikey, id = FALSE)))

tidyverse_rl %>% 
  tidyr::unnest(details)

```
