---
title: "Saving PNG Images"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Saving PNG Images}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r,include=FALSE}
library(httptest)
start_vignette("png")
```

ChemSpider's API services offers the functionality to download image files of the chemicals. This functionality is implemented in the `get_recordId_image()` function. These image files can also be stored straight from R.


### Installing the required `magick` package

Before downloading an example image for caffeine, the `magick` package needs to be installed.
```{r,eval=FALSE}
install.packages("magick")
```


### Downloading the PNG image of caffeine

First, the `chemspiderapi` package needs to be loaded and the API key needs to be retrieved.
```{r,eval=FALSE}
library(chemspiderapi)

apikey <- keyring::key_get(service = "ChemSpider API key", username = Sys.getenv("USERNAME"))
```

Now, the "raw" image file (ChemSpider ID 2424) is downloaded using `get_recordId_image()`.
```{r,eval=FALSE}
caffeine_image_raw <- get_recordId_image(recordId = 2424L, apikey = apikey)
```

Next, the image needs to be read in using the `magick::image_read()` functionality.
```{r,eval=FALSE}
library(magick)

caffeine_png <- image_read(caffeine_image_raw)
```


### Viewing the image

The loaded image can be viewed by calling the object. 
```{r,eval=FALSE}
caffeine_png
```


### Saving the PNG image

Saving the image is possible using the `magick::image_write()` function.
```{r,eval=FALSE}
image_write(caffeine_png, path = "caffeine.png")
```

```{r,include=FALSE}
end_vignette()
```
