---
title: "Storing and Accessing API Keys"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{storing-and-accessing-api-keys}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

> It is good practice to *not* store API keys in code or as plain text!

To store API keys, we recommend using the `keyring` package. This vignette provides a minimal intro to the necessary functionality. For further functionalities of `keyring`, including real keyrings and password protection of the ChemSpider API key, we recommend the documentation of `keyring`, available [here](https://cran.r-project.org/package=keyring).

Before proceeding, the package needs to be installed.
```{r,eval=FALSE}
library("keyring")
```


## Storing an API key using `keyring`

The `keyring::key_set()` will open a pop-up window in which the ChemSpider API key can be pasted in.

```{r,eval=FALSE}
library(keyring)

## Note that Sys.getenv("USERNAME") is used to dynamically tie the API key to the current user.
key_set(service = "ChemSpider API key", username = Sys.getenv("USERNAME"))

```

The name of the stored key is "ChemSpider API key", and the username is set to be the one currently operating R; this can of course be modified. 

## Accessing the API key

The stored API key can be accessed dynamically using the `keyring::key_get()` function. 

```{r,eval=FALSE}
apikey <- key_get(service = "ChemSpider API key", username = Sys.getenv("USERNAME"))
```
