---
title: "Rate-limiting API queries"
author: "Raoul Wolf"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Rate-limiting API queries for `chemspiderapi`

Most users of ChemSpider's API services will have to apply rate limiting conditions. Such a rate limit could be, for example, "no more than 15 queries per minute". The following two paragraph describe how to rate-limit any `chemspiderapi` function; both in `base` R (using `ratelimitr`) and `tidyverse` coding styles.

### A `base` R example

### A `tidyverse` example

In the tidyverse ecosystem, all relevant chemspiderapi functions can be used in functional programming approaches using the purrr::map() family of functions. purrr::slowly() can be used to limit the number of calls of chemspiderapi functions in a given time interval.

```{r,eval=FALSE}
## Installing the necessary packages if necessary
# install.packages(c("chemspiderapi", "tidyverse"))

## Loading the tidyverse package family and chemspiderapi
library(tidyverse)
library(keyring)
library(chemspiderapi)

## Storing the API key as temporary object via the keyring package
## See separate vignette for explanations on how to store security tokens using keyring

apikey <- key_get(service = "ChemSpider API key", username = Sys.getenv("LOGNAME"))

slowly_post_inchikey <- slowly(post_inchikey, rate = rate_delay(pause = 4), quiet = FALSE)

tidyverse_rl <- demo_data %>% 
  select(InChIKey) %>% 
  mutate(queryId = map_chr(InChIKey, ~ slowly_post_inchikey(inchikey = .x, apikey = apikey)))

tidyverse_rl
```

```{r,eval=FALSE}
slowly_get_queryId_status <- slowly(get_queryId_status, rate = rate_delay(pause = 4), quiet = FALSE)

tidyverse_rl <- tidyverse_rl %>% 
  mutate(status = map(queryId, ~ slowly_get_queryId_status(queryId = .x, apikey = apikey)))

tidyverse_rl %>% 
  unnest(status)

tidyverse_rl <- tidyverse_rl %>% 
  unnest(status) %>% 
  select(InChIKey, queryId, status)
```

```{r,eval=FALSE}
slowly_get_queryId_results <- slowly(get_queryId_results, rate = rate_delay(pause = 4), quiet = FALSE)

tidyverse_rl <- tidyverse_rl %>% 
  mutate(id = map2_int(queryId, status, ~ slowly_get_queryId_results(queryId = .x, status = .y, apikey = apikey)))

tidyverse_rl
```

```{r,eval=FALSE}
slowly_get_recordId_details <- slowly(get_recordId_details, rate = rate_delay(pause = 4), quiet = FALSE)

tidyverse_rl <- tidyverse_rl %>% 
  mutate(details = map(id, ~ slowly_get_recordId_details(id = .x, apikey = apikey, id = FALSE)))

tidyverse_rl %>% 
  unnest(details)

```



## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
