---
title: "Saving MOL and SDF Files"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{saving-mol-and-sdf-files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

MOL files are returned by multiple possible API queries. Most obviously by `get_recordId_mol()`, but also as part of `get_recordId_details()`. Saving those human-readable MOL files from R is possible with `base` R functionality.

### Downloading the MOL file of caffeine

For demonstration purposes, we will work with the MOL file for caffeine. Before starting, we need to load the `chemspiderapi` package and retrieve the API key. See the separate vignette "Storing and Accessing API Keys" for details on handling API keys using the `keyring` package.

```{r,eval=FALSE}
library(chemspiderapi)

apikey <- keyring::key_get(service = "ChemSpider API key", username = Sys.getenv("USERNAME"))
```

We can now get the MOL file for caffeine, which has a ChemSpider ID of 2424.

```{r,eval=FALSE}
caffeine_mol <- get_recordId_mol(recordId = 2424, apikey = apikey)
```


### Saving the MOL file

Saving the file is done using the `writeLines()` function. 

```{r,eval=FALSE}
## Writing the MOL file. Note that the file is specified as a connection.
con <- file("caffeine.mol")

writeLines(caffeine_mol, con = con)

close(con)
```

### Saving SDF files

Similar to the above procedure, bulk files SDF files containing multiple MOL entries can be stored. SDF files would be the result of queries by, for example, `post_mass()`. Accordingly, only the file connection needs to be changed. The below example is looking at chemicals

```{r,eval=FALSE}
## First the queryId is obtained
queryId <- post_mass(mass = 194, range = 0.002, apikey = apikey)

## Next, the status is controlled
status <- get_queryId_status(queryId = queryId, count = FALSE, message = FALSE, apikey = apikey)

## Note that per default, the MOL file is decompressed.
caffeine_sdf <- get_queryId_results_sdf(queryId = queryId, status = status, apikey = apikey)

## The file can now be written.
con <- file("caffeine.sdf")

writeLines(caffeine_sdf, con = con)

close(con)
```

